---
title: "ADA_Practical_2"
author: "Zane Hassoun"
date: "2/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(car)
library(lawstat)
library(readr)
HornsRev <- read_csv("C:/Users/zaneh/Downloads/HornsRev.csv")
HornsRev$Impact = as.factor(HornsRev$Impact)
HornsRev$AbundanceRate = HornsRev$Nhat/HornsRev$Area 
```

#### Part I Fitting the Initial Model glmFit0D2 and glmFit0D3 with and without interaction

```{r, echo = FALSE}
glmFit0D2 = glm(Nhat ~ Impact + Depth + XPos + YPos, 
                   data = HornsRev,
                   family = quasipoisson(),
                   offset = log(Area))

glmFit0D3 = glm(Nhat ~ Impact + Depth + XPos + YPos + Impact:YPos + Impact:XPos,
                   data = HornsRev,
                   family = quasipoisson,
                   offset = log(Area))
glmFit0D2
glmFit0D3
```

#### Part II Detecting Collinearity - Pairwise scatter plot using the pairs function

```{r, Echo = FALSE}

covariates = c("XPos", 
                "YPos", 
                "Depth")
pairs(subset(HornsRev, 
             select=covariates),
      upper.panel=NULL, 
      pch=19,
      cex=0.3)

# Now checking the Variance Inflation Factors of the Covariates within Each Model
round(vif(glmFit0D2),2)
round(vif(glmFit0D3),2)
```

#### Extract the model Matrix Itself and now evaluate the plot

```{r, Echo = FALSE}
xmatrix = model.matrix(glmFit0D3)
# Plot the pairs matrix or coliniarity
#head(xmatrix)
covariates <- c("XPos", "YPos", "Depth", "Impact1",
"Impact1:XPos", "Impact1:YPos")
pairs(subset(xmatrix, select=covariates),
upper.panel=NULL, pch=19, cex=0.3)
```

#### Part III Detecting Residual Autocorrelation

```{r, echo = FALSE}
#Generating 50 values from a normal distribution for the use of the model
set.seed(101)
testVals = rnorm(50)
plot(sign(testVals),
     type = "line")
plot(sign(residuals(glmFit0D3, 
                    type = 'pearson')[1:800]),
     type = 'line', 
     ylab = "Sign of the residuals")
#Running a runs test to evaluate the statistics 

runs.test(residuals(glmFit0D3, type = 'pearson'))
```

#### Part IV Diagnosing Non LInearities on the link scale

```{r, echo = FALSE}
par(mfrow=c(2,3))
#Plotting the residual plots in relation to depth
residualPlots(glmFit0D3,
              type="pearson",
              terms=~Depth,
              quadratic=TRUE,
              smooth=list(smoother=gamLine, col="#377eb8"),
              fitted=FALSE,
              col.quad="#e41a1c",
              col="grey",
              pch=19,
              cex=0.3,
              ylim=c(-20, 20))
#Now test compare the residuals positive and negative of runs for first 100 with a random amount
set.seed(101)
# par(mfrow=c(1, 2))
plot(sign(residuals(glmFit0D3, 
                    type="pearson")[order(HornsRev$Depth)])[1:100],
     type="l", 
     main="100 Pearson residuals in depth order",
     ylab="Pearson residuals")

plot(sign(rnorm(100)), type="l", main="Random values")
# The runs values look relatively different as there are longer runs for depth at negative values as opposed to positives
#testing the runs test now to evaluate
runs.test(residuals(glmFit0D3,
                    type = "pearson")[order(HornsRev$Depth)])

#Same exercise but for xpos term
residualPlots(glmFit0D3,
              type="pearson",
              terms=~XPos,
              quadratic=TRUE,
              smooth=list(smoother=gamLine, col="#377eb8"),
              fitted=FALSE,
              col.quad="#e41a1c",
              col="grey",
              pch=19,
              cex=0.3,
              ylim=c(-20, 20))
#Runs test on that data 
runs.test(residuals(glmFit0D3,
                    type = "pearson")[order(HornsRev$XPos)])

#Same exercise for YPos
residualPlots(glmFit0D3,
              type="pearson",
              terms=~YPos,
              quadratic=TRUE,
              smooth=list(smoother=gamLine, col="#377eb8"),
              fitted=FALSE,
              col.quad="#e41a1c",
              col="grey",
              pch=19,
              cex=0.3,
              ylim=c(-20, 20))
#Runs test on that data 
runs.test(residuals(glmFit0D3,
                    type = "pearson")[order(HornsRev$YPos)])

```

```{r, include = FALSE}
#1. A - False, B - False, C - False, D - *True*, E - False D
#2. A - False, B - *True*, C - False, D - False, E - False B
#3. A - False, B - False, C - False, D - False, E - *True* E
#4. A - *True*, B - False, C - False, D - False, E - False A
#5. A - False, B - False, C - False, D - *True*, E - False D
```

#### Part 5 Questions and Answers

1.  **D**
2.  **B**
3.  **E**
4.  **A**
5.  **D**
6.  **-66.63**
7.  **True**
8.  **-35.72**
9.  **True**
10. **-6.29**
11. **True**
12. **True**
