---
title: "ADA_P6"
author: "Zane Hassoun"
date: "3/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mgcv)
library(fields)
df = read.csv("C:/Users/zaneh/Downloads/HornsRev.csv")
pred_data = read.csv("C:/Users/zaneh/Downloads/HornsRevPredictionData.csv")
df$Impact = as.factor(df$Impact)
```

```{r}
colnames(df)
```
## Fitting Two Dimensional Smoothers

1. Produce pre/post-impact spatial plots (using fields::quilt.plot(.)) of the observed data.

```{r, fig.width=8, fig.height=4}
#Fitting the Quilt Plots

par(mfrow=c(1,2))

# Loop across all phases


for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       df$Nhat[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       #zlim=range(NhatPredLink$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact", "Post Impact")))
}
```

2. Fit a quasi-Poisson penalised spline based GAM (mgcv::gam(.); use the default options) with a two
dimensional smoother for the spatial coordinates and a one dimensional smoother for Depth, whilst
including Impact (not as an interaction term for now). Print out a summary of the resultant model,
which will be referred to as PRS_2D.

```{r}
#Fitting a Quasi-Poisson with 2D Smoother for X and Y and 1D for depth and nothing for impact

PRS_2D = mgcv::gam(Nhat ~ s(XPos, YPos) + s(Depth) + Impact,
                   family = quasipoisson,
                   offset = log(Area),
                   data = df)
summary(PRS_2D)
```
3. View the partial plots for PRS_2D.

```{r}
#Viewing the Partial Plots

plot(PRS_2D, shade = T)

```

4. Use HornsRevPredictionData.csv to make pre/post-impact predictions on this fine grid using PRS_2D.
Produce a spatial plot of the results (both on the link and response scale; set the same 𝑧-scale range
(zlim) for pre and post-impact).

```{r}
pred_link = predict(PRS_2D,
                    newdata = pred_data,
                    type = 'link')
pred_probs = predict(PRS_2D,
                    newdata = pred_data,
                    type = 'probs')



plot.new()
# Loop across all phases



for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       pred_link[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                     zlim = range(pred_link),
                       main=paste(ifelse(phase == 0, "Pre Impact (Link Scale)", "Post Impact (Link Scale)")))
}
for(phase in c(0,1)){
  iteration = df$Impact %in% phase
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       pred_probs[which(df$Impact == phase)],
                       nrow=25, ncol=60,zlim = range(pred_probs),
                       main=paste(ifelse(phase == 0, "Pre Impact(Response Scale)", "Post Impact (Response Scale)")))
}


```
## Fitting two dimensional smoothers with an interaction term

5. Fit a similar model to PRS_2D, but now add Impact as an interaction term in the two dimensional
spatial smoother (leave everything else the same). Print out a summary of the resultant model, which
will be referred to as PRS_2DInt
```{r}
#Fitting a Quasi-Poisson with 2D Smoother for X and Y and 1D for depth and nothing for impact

PRS_2DInt = mgcv::gam(Nhat ~ s(XPos, YPos, by = Impact) + s(Depth) + Impact,
                   family = quasipoisson,
                   offset = log(Area),
                   data = df)
summary(PRS_2DInt)
```
6. View the partial plots for PRS_2DInt.

```{r}
#Viewing the Partial Plots

plot(PRS_2DInt, shade = T)

```

7. Use HornsRevPredictionData.csv to make pre/post-impact predictions on this fine grid using PRS_2D.
Produce a spatial plot of the results (both on the link and response scale; set the same 𝑧-scale range
(zlim) for pre and post-impact).

```{r}
pred_linkINT = predict(PRS_2DInt,
                    newdata = pred_data,
                    type = 'link')
pred_probsINT = predict(PRS_2DInt,
                    newdata = pred_data,
                    type = 'probs')



plot.new()
# Loop across all phases



for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       pred_linkINT[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                     zlim = range(pred_linkINT),
                       main=paste(ifelse(phase == 0, "Pre Impact (Link Scale)", "Post Impact (Link Scale)")))
}
for(phase in c(0,1)){
  iteration = df$Impact %in% phase
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       pred_probsINT[which(df$Impact == phase)],
                       nrow=25, ncol=60,zlim = range(pred_probsINT),
                       main=paste(ifelse(phase == 0, "Pre Impact(Response Scale)", "Post Impact (Response Scale)")))
}

```

8. Plot the locations of the observation points and the fine prediction grid (HornsRevPredictionData.csv).

```{r}
library(MRSea)
knotGrid = MRSea::getKnotgrid(coordData = cbind(df$XPos,df$YPos),
                              numKnots = 300,
                              plot = FALSE)

plot(df$XPos, df$YPos, col="grey", pch=19, cex=0.2)
points(knotGrid[, 1], knotGrid[, 2], pch=17, col="#e41a1c")
```
9. Try increasing (set k=15) and decreasing (set k=40) the limit on the degrees of freedom for the two
dimensional spatial smoother (leave everything else the same; keep the interaction term). Print out a
summary of the resultant models.


```{r}
PRS_2DInt_knots15 = mgcv::gam(Nhat ~ s(XPos, YPos, by = Impact, k=15) + s(Depth) + Impact,
                   family = quasipoisson,
                   offset = log(Area),
                   data = df)
summary(PRS_2DInt_knots15)

PRS_2DInt_knots40 = mgcv::gam(Nhat ~ s(XPos, YPos, by = Impact, k=40) + s(Depth) + Impact,
                   family = quasipoisson,
                   offset = log(Area),
                   data = df)
summary(PRS_2DInt_knots40)
```
10. a) for k = 15

```{r}
pred_link15 = predict(PRS_2DInt_knots15,
                    newdata = pred_data,
                    type = 'link')
pred_probs15 = predict(PRS_2DInt_knots15,
                    newdata = pred_data,
                    type = 'probs')



plot.new()
# Loop across all phases



for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       pred_link15[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                     zlim = range(pred_link15),
                       main=paste(ifelse(phase == 0, "Pre Impact (Link Scale)", "Post Impact (Link Scale)")))
}
for(phase in c(0,1)){
  iteration = df$Impact %in% phase
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       pred_probs15[which(df$Impact == phase)],
                       nrow=25, ncol=60,zlim = range(pred_probs15),
                       main=paste(ifelse(phase == 0, "Pre Impact(Response Scale)", "Post Impact (Response Scale)")))
}

```

10. b) for K = 40

```{r}
pred_link40 = predict(PRS_2DInt_knots40,
                    newdata = pred_data,
                    type = 'link')
pred_probs40 = predict(PRS_2DInt_knots40,
                    newdata = pred_data,
                    type = 'probs')



plot.new()
# Loop across all phases



for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       pred_link40[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                     zlim = range(pred_link40),
                       main=paste(ifelse(phase == 0, "Pre Impact (Link Scale)", "Post Impact (Link Scale)")))
}
for(phase in c(0,1)){
  iteration = df$Impact %in% phase
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       pred_probs40[which(df$Impact == phase)],
                       nrow=25, ncol=60,zlim = range(pred_probs40),
                       main=paste(ifelse(phase == 0, "Pre Impact(Response Scale)", "Post Impact (Response Scale)")))
}

```






