---
title: "ADA_P5"
author: "Zane Hassoun"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries
library(tidyverse) # ggplot(.)
library(mgcv) # gam(.)
library(splines) # bs(.) (B-splines)
library(MuMIn) # dredge(.)
library(fields) # quilt.plot(.)
library(lawstat) # runs.test(.)
#Read the Dataset in
df = read.csv("C:/Users/zaneh/Downloads/HornsRev.csv")
# Set Impact as factor
df$Impact = as.factor(df$Impact)
```

#### Penalised Regression Splines using many covariates

1. Fit a quasi-Poisson penalised spline based GAM (mgcv::gam(.)) as follows. Print out a summary of
the model and inspect the results
```{r}
#Penalised Regression Spline

PRS = mgcv::gam(Nhat ~ s(XPos) + s(YPos) + s(Depth) + Impact,
data=df, family=quasipoisson, offset=log(Area))

summary(PRS)

par(mfrow=c(2,2))
plot(PRS, shade = T, residuals = T, ylim=c(-10,10))

```

3. Visually compare the Depth relationship obtained using the PRS model to a model that uses a B-spline
basis function instead with just one knot at 20m

```{r}
PRS_B = stats::update(PRS, .~. -s(Depth) + splines::bs(Depth, knots = 20))
summary(PRS_B)
stats::termplot(PRS_B, se=T)


```


4. Carry out model selection on the terms in the working GAM (i.e. the PRS model fitted in 1.) using the
dredge function in the MuMIn library. Use QAIC to rank the different models. See lecture notes Section
5.5 for a reminder of how to use the dredge function (the overdispersion parameter estimate can be
found in PRS$scale).



```{r}
library(MuMIn)
options(na.action="na.fail") # fail-safe
PRS_Q = update(PRS, family = "poisson")
head(dredge(PRS_Q,
            rank = "QAIC",
            chat = PRS$scale))

```

5. Download HornsRevPredictionData.csv from Moodle. This file contains a fine grid of XPos/YPos
values and corresponding Depth measurements. Use the “best” model identified in 4. to predict over
this fine grid and produce a spatial plot of all the results (pre/post-impact both on the link and response
scale).

```{r}
predData = read.csv("C:/Users/zaneh/Downloads/HornsRevPredictionData.csv")

# Predict on the link scale
NhatPredLink = predict(PRS, newdata=predData,
se=T, type="link")
# Predict on the response scale
NhatPredRes = predict(PRS, newdata=predData,
se=T, type="response")

```
Predictions Quilt Plots

```{r}
par(mfrow=c(2,2))

# Loop across all phases


for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       NhatPredLink$fit[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       #zlim=range(NhatPredLink$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact (Link Scale)", "Post Impact (Link Scale)")))
}
for(phase in c(0,1)){
  iteration = df$Impact %in% phase
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       NhatPredRes$fit[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       # zlim=range(NhatPredRes$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact(Response Scale)", "Post Impact (Response Scale)")))
}

mtext("Quilt Plots for GAM", side = 3, line = -13, outer = TRUE)


```



#### Fitting a GAM with multiple Covariates 

```{r}
PRS_Int = mgcv::gam(Nhat ~ s(XPos, by=Impact) + s(YPos) + s(Depth) + Impact,
                    data=df, 
                    family=quasipoisson, 
                    offset=log(Area))
summary(PRS_Int)

NhatPredLink_2 = predict(PRS_Int, newdata=predData,
se=T, type="link")
# Predict on the response scale
NhatPredRes_2 = predict(PRS_Int, newdata=predData,
se=T, type="response")

```

```{r}
par(mfrow=c(2,2))

# Loop across all phases


for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       NhatPredLink_2$fit[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       zlim=range(NhatPredLink_2$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact (Link Scale)", "Post Impact (Link Scale)")))
}
for(phase in c(0,1)){
  iteration = df$Impact %in% phase
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       NhatPredRes_2$fit[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       zlim=range(NhatPredRes_2$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact(Response Scale)", "Post Impact (Response Scale)")))
}

mtext("Quilt Plots for GAM with Interaction of XPos and Impact", side = 3, line = -13, outer = TRUE)

runs.test(residuals(PRS_Int, type = 'pearson'))
residuals(PRS_Int, type = 'pearson')[1:100]
acf(residuals(PRS_Int, type = 'pearson'))
```

```{r}


par(mfrow=c(2,2))

# Loop across all phases


for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       fitted(PRS)[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       #zlim=range(NhatPredLink$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact (Link Scale)", "Post Impact (Link Scale)")))
}
for(phase in c(0,1)){
  iteration = df$Impact %in% phase
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       fitted(PRS)[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       # zlim=range(NhatPredRes$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact(Response Scale)", "Post Impact (Response Scale)")))
}



par(mfrow=c(2,2))

# Loop across all phases


for(phase in c(0,1)){
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       fitted(PRS_Int)[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       #zlim=range(NhatPredLink$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact (Link Scale)", "Post Impact (Link Scale)")))
}
for(phase in c(0,1)){
  iteration = df$Impact %in% phase
  fields::quilt.plot(df$XPos[which(df$Impact == phase)], 
                       df$YPos[which(df$Impact == phase)],
                       fitted(PRS_Int)[which(df$Impact == phase)],
                       nrow=25, ncol=60,
                       # zlim=range(NhatPredRes$fit),
                       main=paste(ifelse(phase == 0, "Pre Impact(Response Scale)", "Post Impact (Response Scale)")))
}



```

```{r}
q1 = "A"
q2 = "E"
q3 = "True"
q4 = "False"
q5 = "True"
q6 = "C"
q7 = "True"
q8 = "True"
q9 = "False"
q10 = "B"

answers = data.frame("Answer" = c(1:10),
                     "Question" = c(q1,q2,q3,q4,q5,q6,q7,q8,q9,q10)
                     )

pander::pander(answers)
```














